### 2016-06-12 Paper_v4.tex
Updated robustness figure
Introduction is done

The complement cascade involves over 30 soluble and cell surface proteins, receptors and regulators.
The molecular connectivity of complement is complex, see the review of Walport \cite{Walport:2001ab,Walport:2001aa}.
The central outputs of complement are the Membrane Attack Complex (MAC), and the inflammatory mediator proteins C3a and C5a.
The membrane attack complex, generated during the terminal phase of the response,
forms transmembrane channels which disrupt the membrane integrity of targeted cells, leading to cell lysis and death.
On the other hand, the C3a and C5a proteins act as a bridge between innate and adaptive immunity, and play an important role in regulating inflammation \cite{sarma2011complement}.
Complement activation takes places through three pathways: the classical, the lectin binding and the alternate pathways.
Each of these pathways involves a different initiator signal which triggers downstream events in the complement system.
The classical pathway is triggered by antibody recognition of foreign antigens or other pathogens.
A multimeric protein complex C1 binds to antibody-antigen complexes and undergoes a conformational change, leading to an activated form with proteolytic activity.
This activated complex then cleaves soluble complement proteins C4 and C2 into C4a, C4b, C2a and C2b, respectively.
The C4a and C2b fragments bind to form the C4bC2a protease, which is also known as the classical C3 convertase.
The lectin pathway is initiated through the binding of L-ficolin or Mannose Binding Lectin (MBL) to carbohydrates on the surfaces of bacterial pathogens.
These complexes, in combination with the associated mannose-associated serine proteases
1 and 2 (MASP-1/2), also cleave C4 and C2, leading to additional classical C3 convertase.
Thus, the classical and lectin pathways, initiated by the recognition of a foreign surface,  converge at the classical C3 convertase.
However, the alternate pathway works differently.
The alternate pathway involves a 'tickover' mechanism in which complement protein C3 is spontaneously hydrolyzed to form an activated intermediate C3w;
C3w recruits factor B and factor D, leading to the formation of C3wBb. C3wBb can cleave C3 into C3a and C3b, where
the C3b fragment can further recruit additional factor B and factor D to form C3bBbC3b, which is also known as the alternate C3 convertase \cite{pangburn1984alternative}.
The role of classical and alternate C3 convertases is varied.
First, C3 convertases encode an amplification loop by cleaving C3 into C3a and C3b; the C3b fragment is then free to form additional alternate C3 convertases,
thereby forming a positive feedback loop.
Next, C3 convertase activity links complement initiation with the terminal phase of the cascade through the formation of C5 convertases.
Both classical and alternate C3 convertases can recruit an additional C3b subunit to form the classical C5 convertase (C4bC2aC3b), and the alternate C5 convertase (C3bBbC3b), respectively.
C5 convertases cleave C5 into the C5a and C5b fragments.
The C5b fragment, along with the C6, C7, C8 and multiple C9 complement proteins, form the membrane attack complex.
On the other hand, both C3a and C5a are important inflammatory signals involved in several responses \cite{Walport:2001ab,Walport:2001aa}.

%Complement activation is largely irreversible, involving positive feedback driven proteolytic cleavage.
%Thus,
Activation of the complement cascade is strongly regulated by many plasma and host cell proteins.
The initiation of the classical pathway via complement protein C1 is controlled by the C1 Inhibitor (C1-Inh), a protease inhibitor belonging to the serpin superfamily.
C1-Inh irreversibly binds to and deactivates the active subunits of C1, preventing spontaneous fluid phase and chronic activation of complement \cite{walker1995complement}.
Regulation of the upstream elements of complement is also achieved through the interaction of the C4 binding protein (C4BP) with C4b, as well as
through the interaction of factor H with C3b \cite{blom2001structural}.
These regulatory proteins are also capable of binding their respective targets while they are bound in convertase complexes.
Membrane cofactor protein (MCP or CD46) possesses a cofactor activity for C4b and C3b, which protects the host from self-activation of complement \cite{riley2004cd46}.
Delay accelerating factor (DAF or CD55) is also able to recognize and dissociate both C3 and C5 convertases \cite{lukacik2004complement}.
Carboxypeptidase-N, a well known inflammation regulator, cleaves carboxyl-terminal arginines and lysines of the complement proteins C3a, C4a, and C5a rendering them inactive \cite{liszewski1995control}. Lastly, the assembly of the MAC complex is inhibited by vitronectin and clusterin in the plasma, and CD59 at the host surface \cite{chauhan2006presence,zewde2016quantitative}.
Thus, there are many points of control which influence complement activation across the three activation pathways.

Developing quantitative mathematical models of complement could be crucial to understanding its role in the body.
Traditionally, complement models have been formulated as systems of linear or non-linear ordinary differential equations (ODEs).
For example, Hirayama et al. modeled the classical complement pathway as a system of linear ODEs \cite{hirayama1996linear},
while Korotaevskiy and co-workers modeled the classical, lectin and alternate pathways as a system of non-linear ODEs \cite{korotaevskiy2009non}.
More recently, large mechanistic models of sections of complement have also been proposed.
For example, Liu et al. analyzed the formation of the classical and lectin C3 convertases, and the regulatory role of C4BP using a system of 45 non-linear ODEs with 85 parameters \cite{liu2011computational}.
Recently, Zewde and co-workers constructed a detailed mechanistic model of the alternative pathway which consisted of 107 ODEs and 74 kinetic parameters and delineated
the complement response of the host and pathogen \cite{zewde2016quantitative}.
However, these previous modeling studies involved little experimental validation.
Thus, while these models are undoubtably important theoretical tools, it is unclear if they can describe or quantitatively predict experimentally validated complement dynamics.
The central challenge is the estimation of model parameters from experimental data.
Unlike other important cascades, such as coagulation for which there are well developed experimental tools and many publicly available data sets,
the data for complement is relatively sparse. Missing or incomplete data sets, and limited quantitative data make the identification of mechanistic complement models difficult.

In this study, we developed an ensemble of experimentally validated reduced order complement models.
The modeling approach combined ordinary differential equations with logical rules to produce a complement model with a limited number of equations and parameters.
The reduced order model,  which described the lectin and alternative pathways, consisted of 18 differential equations with 28 parameters.
Thus, the model was an order of magnitude smaller and included more pathways than comparable mathematical models in the literature.
We estimated an ensemble of model parameters from \textit{in~vitro} time series measurements of the C3a and C5a complement proteins.
Subsequently, we validated the model on unseen C3a and C5a measurements that were not used for model training.
Despite its small size, the model was surprisingly predictive.
After validation, we performed global sensitivity and robustness analysis to estimate which parameters and species
controlled model performance. These analyses suggested complement was robust to any single therapeutic intervention.
The only intervention that consistently reduced C5a formation for all cases was a dual-knockdown of both C3 and C5.
Taken together, we developed a reduced order complement model that was computationally inexpensive,
and could easily be incorporated into pre-existing or new pharmacokinetic models of immune system function.
The model described experimental data, and predicted the need for multiple points of intervention to disrupt complement activation.
%Taken together, the combined analysis of alternate and lectin pathways along with the incorporation of the downstream reactions involving C5 convertase
%elucidated new insight into the roles of parameters that govern the complement system.  A deeper understanding about how these parameters influence complement dynamics will greatly aid in the development of drugs for strategic therapeutic targets. Due to the low computational cost relative to the existing models and accuracy of our predictions, we believe that our reduced order complement network is the first step towards building a computation toolbox for screening drug potential drug targets or therapeutic agents that can be targeted against complement.

@@

@@

Robustness coefficients quantify the response of a marker to a macroscopic structural or operational perturbation to a biochemical network.
Here, we computed how the C3a and C5a trajectories responded to a decrease in the initial abundance of C3 and/or C5.
Robustness analysis suggested there was no single intervention that inhibited complement activation in the presence of both initiation pathways (Fig. \ref{fig-robustness-analysis}).
We calculated robustness indices for C3a and C5a for the parameter ensemble (N = 50) with and without the lectin pathway initiator.
We simulated the addition of different doses of anti-complement inhibitor cocktails by decreasing the initial concentration of C3 or C5 or the combination of C3 and C5 by 50\% and 90\%.
This would be conceptually analogous to the administration of a C3 inhibitor e.g., compstatin alone or combination with eculizumab (Fig. \ref{fig-SA}E).
A \texttt{log$_{10}$} transformed robustness index of zero indicated no effect due to the perturbation, whereas an index of less than zero indicated decreased C3a or C5a.
The response of the complement model to different knock-down magnitudes was non-linear; a 90\% knock-down had an order of magnitude more impact than a 50\% knock-down.
As expected, a C5 knockdown had no effect on C3a formation for either the alternate (Fig. \ref{fig-robustness-analysis}A, lanes 1 or 3) or lectin pathways
(Fig. \ref{fig-robustness-analysis}B, lanes 1 or 3).
However, C3a abundance and to a lesser extent C5a abundance decreased with decreasing C3 concentration in the alternate pathway (Fig. \ref{fig-robustness-analysis}A or B, lanes 1 or 2).
This agreed with the sensitivity results; changes in AP C3-convertase formation or activity affected the downstream dynamics of C5a formation.
Thus, if we only considered the alternate pathway, C3 alone could be a reasonable target, especially given that C5a formation was surprisingly robust to C5 levels in the alternate pathway (Fig. \ref{fig-robustness-analysis}A or B, lane 2).
Yet, when both pathways were activated, C5a levels were robust to the initial C3 concentration (Fig. \ref{fig-robustness-analysis}A or B, lane 4);
C5a formation was catalyzed by CP C3 and C5 convertases.
Thus, the only reliable intervention that consistently reduced both C3a and C5a formation for all cases was a dual-knockdown of C3 and C5.
For example, a 90\% decrease of both C3 and C5 reduced the formation of C5a by over an order of magnitude (Fig. \ref{fig-robustness-analysis}B, lane 4),
while C3a was reduced to a lesser extent (Fig. \ref{fig-robustness-analysis}B, lane 3).

>>>>

The complement cascade involves over 30 soluble and cell surface proteins, receptors and regulators.
The molecular connectivity of complement involves both fluid phase and cell surface events, see the review of Walport \cite{Walport:2001ab,Walport:2001aa}.
The central outputs of complement are the Membrane Attack Complex (MAC), and the inflammatory mediator proteins C3a and C5a.
The membrane attack complex, generated during the terminal phase of the response,
forms transmembrane channels which disrupt the membrane integrity of targeted cells, leading to cell lysis and death.
On the other hand, the C3a and C5a proteins act as a bridge between innate and adaptive immunity, and play an important role in regulating inflammation \cite{sarma2011complement}.
Complement activation takes places through three pathways: the classical, the lectin binding and the alternate pathways.
The classical pathway is triggered by antibody recognition of foreign antigens or other pathogens.
A multimeric protein complex C1 binds to antibody-antigen complexes and undergoes a conformational change, leading to an activated form with proteolytic activity.
This activated complex then cleaves soluble complement proteins C4 and C2 into C4a, C4b, C2a and C2b, respectively.
The C4a and C2b fragments bind to form the C4bC2a protease, also known as the classical C3 convertase.
The lectin pathway is initiated through the binding of L-ficolin or Mannose Binding Lectin (MBL) to carbohydrates on the surfaces of bacterial pathogens.
These complexes, in combination with the associated mannose-associated serine proteases
1 and 2 (MASP-1/2), also cleave C4 and C2, leading to additional classical C3 convertase.
Thus, the classical and lectin pathways, initiated by the recognition of different cues on foreign surfaces,  converge at the classical C3 convertase.
However, the alternate pathway works differently. It is activated by a 'tickover' mechanism in which complement protein C3 is spontaneously hydrolyzed to form an activated intermediate C3w;
C3w recruits factor B and factor D, leading to the formation of C3wBb. C3wBb cleaves C3 into C3a and C3b, where
the C3b fragment further recruits additional factor B and factor D to form C3bBbC3b, the alternate C3 convertase \cite{pangburn1984alternative}.
The role of classical and alternate C3 convertases is varied.
First, C3 convertases mediate signal amplification.
C3 convertases cleave C3 into C3a and C3b; the C3b fragment is then free to form additional alternate C3 convertases,
thereby forming a positive feedback loop.
Next, C3 convertases link complement initiation with the terminal phase of the cascade through the formation of C5 convertases.
Both classical and alternate C3 convertases can recruit C3b subunits to form the classical C5 convertase (C4bC2aC3b), and the alternate C5 convertase (C3bBbC3b), respectively.
C5 convertases cleave C5 into the C5a and C5b fragments.
The C5b fragment, along with the C6, C7, C8 and multiple C9 complement proteins, form the membrane attack complex.
On the other hand, both C3a and C5a are important inflammatory signals involved in several responses \cite{Walport:2001ab,Walport:2001aa}.
Thus, the complement cascade directly attacks the invading pathogen, and signals for help from the immune system.

%Complement activation is largely irreversible, involving positive feedback driven proteolytic cleavage.
%Thus,
The complement cascade is regulated by plasma and host cell proteins which balance host cell safety with effectiveness.
The initiation of the classical pathway via complement protein C1 is controlled by the C1 Inhibitor (C1-Inh).
C1-Inh irreversibly binds to and deactivates the active subunits of C1, preventing chronic complement activation \cite{walker1995complement}.
Regulation of upstream processes in the lectin and alternate pathways also occurs through the interaction of the C4 binding protein (C4BP) with C4b, and factor H with C3b \cite{blom2001structural}. Interestingly,
both factor H and C4BP are capable of binding their respective targets while in convertase complexes as well.
At the host cell surface, membrane cofactor protein (MCP or CD46) can interact with C4b and C3b, which protects the host cell from complement self-activation \cite{riley2004cd46}.
Delay accelerating factor (DAF or CD55) also recognizes and dissociates both C3 and C5 convertases on host cell surfaces \cite{lukacik2004complement}.
More generally, Carboxypeptidase-N, a well known inflammation regulator, has broad activity against the complement proteins C3a, C4a, and C5a, rendering them inactive by cleavage of carboxyl-terminal arginine and lysine residues \cite{liszewski1995control}.
Although Carboxypeptidase-N does not directly influence complement activation, it silences the important inflammatory signals produced by complement.
Lastly, the assembly of the MAC complex itself can be inhibited by vitronectin and clusterin in the plasma, and CD59 at the host surface \cite{chauhan2006presence,zewde2016quantitative}.
Thus, there are many points of control which influence complement across the three activation pathways.

Developing quantitative mathematical models of complement will be crucial to fully understanding its role in the body.
Traditionally, complement models have been formulated as systems of linear or non-linear ordinary differential equations (ODEs).
For example, Hirayama et al. modeled the classical complement pathway as a system of linear ODEs \cite{hirayama1996linear},
while Korotaevskiy and co-workers modeled the classical, lectin and alternate pathways (up to C3a) as a system of non-linear ODEs \cite{korotaevskiy2009non}.
More recently, large mechanistic models of sections of complement have also been proposed.
For example, Liu et al. analyzed the formation of the classical and lectin C3 convertases, and the regulatory role of C4BP using a system of 45 non-linear ODEs with 85 parameters \cite{liu2011computational}.
Zewde and co-workers constructed a detailed mechanistic model of the alternative pathway which consisted of 107 ODEs and 74 kinetic parameters and delineated between
the host cell and pathogen \cite{zewde2016quantitative}.
However, these previous modeling studies involved little experimental validation.
Thus, while these models are undoubtably important theoretical tools, it is unclear if they can describe or quantitatively predict complement measurements.
The central challenge of complement model identification is the estimation of model parameters from experimental measurements.
Unlike other important cascades, such as coagulation where there are well developed experimental tools and publicly available data sets,
the data for complement is relatively sparse. Missing or incomplete data sets, and limited quantitative data make the identification of large mechanistic complement models difficult.
Thus, reduced order modeling approaches, which describe the key biology with a limited number of parameters, will be important for developing validated models of complement, and to fully understandings its role in the body.

%Taken together, the combined analysis of alternate and lectin pathways along with the incorporation of the downstream reactions involving C5 convertase
%elucidated new insight into the roles of parameters that govern the complement system.  A deeper understanding about how these parameters influence complement dynamics will greatly aid in the development of drugs for strategic therapeutic targets. Due to the low computational cost relative to the existing models and accuracy of our predictions, we believe that our reduced order complement network is the first step towards building a computation toolbox for screening drug potential drug targets or therapeutic agents that can be targeted against complement.

@@

In this study, we developed an ensemble of experimentally validated reduced order complement models.
The modeling approach combined ordinary differential equations with logical rules to produce a complement model with a limited number of equations and parameters.
The reduced order model,  which described the lectin and alternative pathways, consisted of 18 differential equations with 28 parameters.
Thus, the model was an order of magnitude smaller and included more pathways than comparable models in the literature.
We estimated an ensemble of model parameters from \textit{in~vitro} time series measurements of the C3a and C5a complement proteins.
Subsequently, we validated the model on unseen C3a and C5a measurements that were not used for model training.
Despite its small size, the model was surprisingly predictive.
After validation, we performed global sensitivity and robustness analysis to estimate which parameters and species
controlled model performance. These analyses suggested complement was robust to any single therapeutic intervention;
only the knockdown of both C3 and C5 consistently reduced C3a and C5a formation for all cases.
Taken together, we developed a reduced order complement model that was computationally inexpensive,
and could easily be incorporated into pre-existing or new pharmacokinetic models of immune system function.
The model described experimental data, and predicted the need for multiple points of intervention to disrupt complement activation.

@@

Robustness coefficients quantify the response of a marker to a macroscopic structural or operational perturbation to a biochemical network.
Here, we computed how the C3a and C5a trajectories responded to a decrease in the initial abundance of C3 and/or C5.
Robustness analysis suggested there was no single intervention that inhibited complement activation in the presence of both initiation pathways (Fig. \ref{fig-robustness-analysis}).
We calculated robustness indices for C3a and C5a for the parameter ensemble (N = 50) with and without the lectin pathway initiator.
We simulated the addition of different doses of anti-complement inhibitor cocktails by decreasing the initial concentration of C3 or C5 or the combination of C3 and C5 by 50\% and 90\%.
This would be conceptually analogous to the administration of a C3 inhibitor e.g., compstatin alone or combination with eculizumab (Fig. \ref{fig-SA}E).
A \texttt{log$_{10}$} transformed robustness index of zero indicated no effect due to the perturbation, whereas an index of less than zero indicated decreased C3a or C5a.
The response of the complement model to different knock-down magnitudes was non-linear; a 90\% knock-down had an order of magnitude more impact than a 50\% knock-down.
As expected, a C5 knockdown had no effect on C3a formation for either the alternate (Fig. \ref{fig-robustness-analysis}A, lanes 1 or 3) or lectin pathways
(Fig. \ref{fig-robustness-analysis}B, lanes 1 or 3).
However, C3a abundance and to a lesser extent C5a abundance decreased with decreasing C3 concentration in the alternate pathway (Fig. \ref{fig-robustness-analysis}A or B, lanes 1 or 2).
This agreed with the sensitivity results; changes in AP C3-convertase formation affected the downstream dynamics of C5a formation.
Thus, if we only considered the alternate pathway, C3 alone could be a reasonable target, especially given that C5a formation was surprisingly robust to C5 levels in the alternate pathway (Fig. \ref{fig-robustness-analysis}A or B, lane 2).
Yet, when both pathways were activated, C5a levels were robust to the initial C3 concentration (Fig. \ref{fig-robustness-analysis}A or B, lane 4);
C5a formation was catalyzed by CP C3 and C5 convertases.
Thus, the only reliable intervention that consistently reduced both C3a and C5a formation for all cases was a dual-knockdown of C3 and C5.
For example, a 90\% decrease of both C3 and C5 reduced the formation of C5a by over an order of magnitude (Fig. \ref{fig-robustness-analysis}B, lane 4),
while C3a was reduced to a lesser extent (Fig. \ref{fig-robustness-analysis}B, lane 3).
